Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-06-07 Master Thesis/synopsis.tex
Fix 3.2 argumentation

\begin{theorem}\label{theorem-absoluteinertia}
Let $K$ be a henselian field with respect to $v: K\to\Gamma\cup\{\infty\}$ with unique extension $\overline{v}$ on $K^\text{s}$. Let $K^\text{t}$ be the absolute inertia field of $K$ with the unique extension $w$ of $v$, i.e. $K^\text{t}$ is the fixed field of the absolute inertia group $T$: 
\[ T = \{ \sigma\in G_K\mid \sigma(x)\equiv x\mod \mathfrak{m}_{\overline{v}} \quad \text{for all }x\in\mathcal{O}_{\overline{v}} \}\lhd G_K \]
Then we have:
\begin{enumerate}[(i)]
\item $w$ has value group $\Gamma$ and residue class field $k(v)^\text{s}$.
\item $K^\text{t}$ is a Galois extension of $K$ and $\operatorname{Gal}(K^\text{t}/K)\cong G_{k(v)}$ as profinite groups. There exists an inclusion-preserving one-to-one correspondence: 
\begin{align*}
\{ \text{Intermediate extensions in $K^\text{t}/K$} \} &\longrightarrow \{\text{Separable extensions of $k(v)$} \}\\ 
L &\longmapsto k(v_L)  
\end{align*}
where $v_L$ is the unique extension of $v$ to $L$.
\end{enumerate}
\end{theorem}

@@

This lemma yields a bijection $\sigma:\operatorname{Spv}_\text{d}^{(2)}(K) \to\operatorname{Spv}_\text{d}^{(2)}(K')$. We will see that this bijection restricts to a bijection $\operatorname{Spv}_\text{rd}^{(2)}(K)\to \operatorname{Spv}_\text{rd}^{(2)}(K')$. With that, we shall establish a bijection between the defectless valuations of rational rank $1$:

@@

\begin{proof}
Lemma \ref{3.1} shows that $\sigma$ induces a bijection $\sigma: \operatorname{Spv}_\text{d}^{(2)}(K)\to \operatorname{Spv}_\text{d}^{(2)}(K')$. For all $v\in\operatorname{Spv}(K)$ we choose a fixed henselization $K_v\subset\overline{K}$ with Lemma \ref{lemma-henselization-specialization}, such that if $v,w\in\operatorname{Spv}(K)$ and $w$ is a specialization of $v$ we have $K_v\subset K_w$. If $v$ is nontrivial, then in this case we get $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ and $v=v(w)\in\operatorname{Spv}_\text{d}^{(1)}(K)$. We set $K'_{\sigma(w)} = (K_w)'$ for $w\in\operatorname{Spv}_\text{d}^{(2)}(K)$. Lemma \ref{3.1} shows that $K'_{\sigma(w)}$ is a henselization of $K'$ with respect to $\sigma(w)$.

Let $w_1,w_2 \in\operatorname{Spv}_\text{d}^{(2)}(K)$ be two different valuations. By Lemma \ref{2.2}, the following statements are equivalent:
\begin{enumerate}[(i)]
\item The composite $K_{w_1}K_{w_2}$ is not algebraically closed.
\item $w_1,w_2\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ and $v(w_1) = v(w_2)$.
\end{enumerate}
If (ii) holds, $k(v)$ is separably closed where $v=v(w_1)$. By Lemma \ref{lemma-henselization-specialization} we have $K_{w_1}K_{w_2}\subset K_v^\text{t}$ and by Theorem \ref{theorem-absoluteinertia} (ii) $K_{w_1}K_{w_2}$ is the absolute inertia field $K_v^\text{t}$ of $K_v$. Furthermore, $K'_{\sigma(w_1)}K'_{\sigma(w_2)} = (K_{w_1}K_{w_2})' \neq \overline{K'}$ and again by Lemma \ref{2.2} we get: \[\sigma(w_1),\sigma(w_2)\in\operatorname{Spv}_\text{rd}^{(2)}(K)\quad \text{and} \quad v(\sigma(w_1)) = v(\sigma(w_2))\]
Thus $\sigma: \operatorname{Spv}_\text{d}^{(2)}(K)\to \operatorname{Spv}_\text{d}^{(2)}(K')$ maps $\operatorname{Spv}_\text{rd}^{(2)}(K)$ onto $\operatorname{Spv}_\text{rd}^{(2)}(K')$ such that:
\[ v(w_1)=v(w_2) \iff v(\sigma(w_1)) = v(\sigma(w_2)) \]
Let $v\in\operatorname{Spv}_\text{d}^{(1)}(K)$. Since $v\not\in |\operatorname{Spv}(K)|$, there exists a $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ such that $v(w)=v$. We define $v'=v(\sigma(w))$. This shows that $\sigma$ also induces a bijection: 
\[\sigma:\operatorname{Spv}_\text{d}^{(1)}(K)\to\operatorname{Spv}_\text{d}^{(1)}(K'),\ v\mapsto v'\] 
This map is well-defined and for all $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ we have $\sigma(v(w)) = v(\sigma(w))$.

Let $v\in\operatorname{Spv}_\text{d}^{(1)}(K)$ and let $N$ be the normal subgroup of $G_{K_v}$ generated by all subgroups $G_{K_w}$ for $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ with $v(w) = v$. Since a global field has no non-trivial algebraic extension in which every prime splits completely, we get $N=G_{K_v}$. This shows that $(K_v)'$ is a henselization of $K'$ with respect to $v'=\sigma(v)$. We write $K'_{v'} = (K_v)'$. Furthermore, let $w_1,w_2\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ with $w_1\neq w_2$ and $v=v(w_1)=v(w_2)$, then:
\[ \sigma(G_{K^\text{t}_v}) = \sigma(G_{K_{w_1}}\cap G_{K_{w_2}}) = G_{K'_{w'_1}}\cap G_{K'_{w'_2}} = G_{(K')^\text{t}_{v'}} \]
Therefore, $\sigma$ induces with Theorem \ref{theorem-absoluteinertia} the following isomorphism:
\[ G_{k(v)} \cong G_{K_v}/G_{K_v^\text{t}} \to G_{K'_{v'}}/G_{(K')_{v'}^\text{t}}\cong G_{k(v')} \]
The rest follows from the Theorem of Neukirch-Uchida \ref{thm:neukirch-uchida}.
\end{proof}

>>>>

\begin{theorem}\label{theorem-absoluteinertia}
Let $K$ be a henselian field with respect to $v: K\to\Gamma\cup\{\infty\}$ with unique extension $\overline{v}$ on $K^\text{s}$. Let $K^\text{t}$ be the absolute inertia field of $K$ with the unique extension $w$ of $v$, i.e. $K^\text{t}$ is the fixed field of the absolute inertia group $T$: 
\[ T = \{ \sigma\in G_K\mid \sigma(x)\equiv x\mod \mathfrak{m}_{\overline{v}} \quad \text{for all }x\in\mathcal{O}_{\overline{v}} \}\lhd G_K \]
Then we have:
\begin{enumerate}[(i)]
\item $w$ has value group $\Gamma$ and residue class field $k(v)^\text{s}$.
\item $K^\text{t}$ is a Galois extension of $K$ and $\operatorname{Gal}(K^\text{t}/K)\cong G_{k(v)}$ as profinite groups. This isomorphism induces an inclusion-preserving one-to-one correspondence: 
\begin{align*}
\{ \text{Intermediate extensions in $K^\text{t}/K$} \} &\longrightarrow \{\text{Separable extensions of $k(v)$} \}\\ 
L &\longmapsto k(v_L)  
\end{align*}
where $v_L$ is the unique extension of $v$ to $L$.
\end{enumerate}
\end{theorem}

@@

This lemma yields a bijection $\sigma:\operatorname{Spv}_\text{d}^{(2)}(K) \to\operatorname{Spv}_\text{d}^{(2)}(K'),\ w\mapsto w'$. We will see that this bijection restricts to a bijection $\operatorname{Spv}_\text{rd}^{(2)}(K)\to \operatorname{Spv}_\text{rd}^{(2)}(K')$. With that, we shall establish a bijection between the defectless valuations of rational rank $1$:

@@

\begin{proof}
Lemma \ref{3.1} shows that $\sigma$ induces a bijection $\sigma: \operatorname{Spv}_\text{d}^{(2)}(K)\to \operatorname{Spv}_\text{d}^{(2)}(K')$. For all $v\in\operatorname{Spv}(K)$ we choose a fixed henselization $K_v\subset\overline{K}$ with Lemma \ref{lemma-henselization-specialization}, such that if $v,w\in\operatorname{Spv}(K)$ and $w$ is a specialization of $v$ we have $K_v\subset K_w$. If $v$ is nontrivial, then in this case we get $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ and $v=v(w)\in\operatorname{Spv}_\text{d}^{(1)}(K)$. We set $K'_{w'} = (K_w)'$ for $w\in\operatorname{Spv}_\text{d}^{(2)}(K)$. Lemma \ref{3.1} shows that $K'_{w'}$ is a henselization of $K'$ with respect to $w'$.

Let $w_1,w_2 \in\operatorname{Spv}_\text{d}^{(2)}(K)$ be two different valuations. By Lemma \ref{2.2}, the following statements are equivalent:
\begin{enumerate}[(i)]
\item The composite $K_{w_1}K_{w_2}$ is not algebraically closed.
\item $w_1,w_2\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ and $v(w_1) = v(w_2)$.
\end{enumerate}
If (ii) holds, $k(v)$ is separably closed where $v=v(w_1)$. By Lemma \ref{lemma-henselization-specialization} we have $K_{w_1}K_{w_2}\subset K_v^\text{t}$ and by Theorem \ref{theorem-absoluteinertia} (ii) $K_{w_1}K_{w_2}$ is the absolute inertia field $K_v^\text{t}$ of $K_v$. Furthermore, $K'_{w'_1}K'_{w'_2} = (K_{w_1}K_{w_2})' \neq \overline{K'}$ and again by Lemma \ref{2.2} we get: \[w'_1,w'_2\in\operatorname{Spv}_\text{rd}^{(2)}(K)\quad \text{and} \quad v(w'_1) = v(w'_2)\]
Thus $\sigma: \operatorname{Spv}_\text{d}^{(2)}(K)\to \operatorname{Spv}_\text{d}^{(2)}(K')$ maps $\operatorname{Spv}_\text{rd}^{(2)}(K)$ onto $\operatorname{Spv}_\text{rd}^{(2)}(K')$ such that:
\[ v(w_1)=v(w_2) \iff v(w'_1) = v(w'_2) \]
Let $v\in\operatorname{Spv}_\text{d}^{(1)}(K)$. Since $v\not\in |\operatorname{Spv}(K)|$, there exists a $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ such that $v(w)=v$. We define $v'=v(w')$. This shows that $\sigma$ also induces a bijection: 
\[\sigma:\operatorname{Spv}_\text{d}^{(1)}(K)\to\operatorname{Spv}_\text{d}^{(1)}(K'),\ v\mapsto v'\] 
This map is well-defined and for all $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ we have $\sigma(v(w)) = v(\sigma(w))$.

Let $v\in\operatorname{Spv}_\text{d}^{(1)}(K)$ and $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ with $v(w)=v$. Then $k(v)$ is finitely generated of dimension $1$ by Lemma \ref{rank-dimension-inequality}, thus a global field. By Theorem \ref{theorem-absoluteinertia}, we have: 
\[G_{K_w}/H \cong \operatorname{Gal}(K_v^\text{t}/K_w)\cong G_{k(v_w)}\] 
where $v_w$ is the unique extension of $v$ on $K_w$ and $H=G_{K_v^\text{t}}$. With Lemma \ref{2.2-pre} we can see that $k(v_w)$ is the henselization of $k(v)$ with respect to $w\mod v\in\operatorname{Spv}k(v)$. Let $N$ be the closed subgroup of $G_{K_v}$ generated by all subgroups $G_{K_w}$ for $w\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ with $v(w) = v$. Since a global field has no non-trivial algebraic extension in which every prime splits completely, we get $N/H= G_{K_v}/H\cong G_{k(v)}$, thus $N=G_{K_v}$. Since $(K_w)'=K'_{w'}$, this shows that $(K_v)'$ is a henselization of $K'$ with respect to $v'=\sigma(v)$. We write $K'_{v'} = (K_v)'$.

Furthermore, let $w_1,w_2\in\operatorname{Spv}_\text{rd}^{(2)}(K)$ with $w_1\neq w_2$ and $v=v(w_1)=v(w_2)$, then:
\[ \sigma(G_{K^\text{t}_v}) = \sigma(G_{K_{w_1}}\cap G_{K_{w_2}}) = G_{K'_{w'_1}}\cap G_{K'_{w'_2}} = G_{(K')^\text{t}_{v'}} \]
Therefore, $\sigma$ induces with Theorem \ref{theorem-absoluteinertia} the following isomorphism:
\[ G_{k(v)} \cong G_{K_v}/G_{K_v^\text{t}} \to G_{K'_{v'}}/G_{(K')_{v'}^\text{t}}\cong G_{k(v')} \]
The rest follows from the Theorem of Neukirch-Uchida \ref{thm:neukirch-uchida}.
\end{proof}
